Put Title here

### 2018-12-12 General WLD combinatorics.tex
recovered lost changes
Made changes to section 1 that had been lost.

%*****************
%Shortcuts for frequently used special characters
\newcommand{\fB}{\mathfrak{B}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\fZ}{\mathfrak{Z}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\Prop}{\textrm{Prop}}
\newcommand{\cW}{\mathcal{W}}
\newcommand{\bM}{\mathbb{M}}
\newcommand{\cZ}{\mathcal{Z}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\Dom}{\textrm{Dom}}
\newcommand{\detzr}[1] {\langle (\cZ_*^\mu|V_p)^{#1} \rangle}
\newcommand{\II}{\mathcal{I}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\BB}{\mathcal{B}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\interval}[2]{[\![#1,#2]\!]}
\newcommand{\gale}[1]{\preccurlyeq_{#1}}
\newcommand{\sgale}[1]{\prec_{#1}}
%*****************

@@

\begin{dfn}\label{WLdfn}
A Wilson loop diagram is given by the following data: a cyclicly ordered set $V$, along with a choice of first vertex, and $k$ pairs $\{p_r = (i_r, j_r)\}_{r=1}^k$ ordered such that $i_r +1 < j_r$. \end{dfn}

We depict this data as a convex polygon, with vertices labeled by $V$ (preserving the cyclic ordering), and $k$ wavy lines in the interior of the diagrams. We depict the choice of first vertex on $V$ by a $\bullet$. \note{If we look at the pictures we have later (at least some of them), we don't do that.  Instead all the vertices are $\bullet$.} Note that the polygon gives the verices of $W$ a cyclic ordering, which the choice of first vertex gives it a compatible linear order. Both the cyclic and the linear order become the correct perspective at various points in this paper. The pair $p_r$ defines a wavy line from the $i_r^{th}$ edge (between $i_r$ and $i_r+1$) and the $j_r^{th}$ edge. These are called propagators.  The condition on $i_r$ and $j_r$ means that the propagator does not go between adjacent edges. Let $\cP = \{p_r\}_{r=1}^k$ be the set of propagators. Then we write \bas W = (\cP, V) \;.\eas

Often we take $V$ to be $[n]$, the cyclically ordered set of integers, $1 \ldots n$. In this case, we write $W = (\cP, n)$. We introduce some notation to speak of vertices supporting a propagator, and the set of propagators supported on a vertex set.

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, n)$.
\begin{enumerate}
\item For $p \in \cP$, let $V_p = \{i_p, i_p+1, j_p, j_p+1\}$ be the set of vertices supporting $p$. Then, for $P \subseteq \cP$, the set $V_P = \cup_{p \in P} V_p$ is the vertex support of $P$.
\item For $V \subseteq [n]$, write $\Prop(V) = \{ p \in \cP | V_p \cap V \neq \emptyset \} $.
\item For $P \subseteq \cP$, define $F(P) = V(P^c)^c$ to be the set of vertices in $[n]$ that only support propagators in the set $P$.
\end{enumerate}
\end{dfn}

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram is admissible if \begin{enumerate}
\item $|V| \geq |\cP| + 4$
\item There does not exists a set of propagators, $P \subset \cP$ such that $|V(P)| < |P| + 3$.
\item There does not exist a pair of propagators, $p, q \subset \cP$ such that $i_p < i_q < j_p <j_q$.
\end{enumerate}
A Wilson loop diagram is weakly admissible if the second and third conditions hold.
 \end{dfn}

The first conditions states that there are at least four more vertices than propagators in an admissible Wilson Loop Diagram. The second imposes an upper bound on how densely the propagators can be fitted in the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram. In other words, a Wilson loop diagram, $(\cP, n)$ is admissible if and only if $n > \cP +4$, and has neither crossing propagators nor any pairs of propagators that start and end on the same pair of non-adjacent edges.

@@

\begin{dfn} \label{subdiagramdfn}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram. A subdiagram of $W$ is defined by a a subset of propagators of $W$. For $P \subset \cP$, write $W|_P = (P, V(P))$.
\end{dfn}
\note{Later I will also want a different kind of subobject, one with a subset of the propagators but where the vertex set contains $V(P)$ but may be larger (the most important instance being when it is $[n]$.)  This kind of subobject doesn't need a particular name, nor does it need to be defined up here, but I wanted to mention it so we have it in mind.}

Note that a subdiagram of an admissible diagram need not be admissible. In particular, the condition $|V(P)| > |P| +4$ need not hold.

@@

\begin{dfn}\label{equivdfn} 
There is an equivalence relationship on the set of admissible Wilson loops diagrams given by the transititve closure the following binary relation: $W = (\cP, n) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

Each Wilson loop diagram, $W = (\cP, n)$ with $|\cP| = k$ is associated to a $k \times n$ matrix with non-zero real variable entries, called $C(W)$:

@@

\begin{dfn}\label{def R(W)}
For $W = (\cP, n)$, $R(W) = \prod_{e=1}^n R_e$, with $R_e$ defined by the propagators ending on it. For any edge $e$ of $W$, order the propagators incident on $e$ as $\{p_1 \ldots p_r\}$, ordered such that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$. Then \bas R_e =  c_{p_1,e+1} \prod_{j= 1}(r-1) \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_s,e}\;.\eas Note that in this notation, if $r = 1$, $R_e = c_{p,e} c_{p,e+1}$.
\end{dfn}

@@

Finally, we describe several important types of matroid. A matroid of rank $k$ with a ground set of size $n$ is called {\em realizable} if there exists some $A \in Gr(k,n)$ whose non-zero $k\times k$ minors are exactly those with columns indexed by elements of $\cB$. A {\em positroid} is a matroid which can be realized by an element of the totally nonnegative Grassmannian $\Gr(k,n)$. Finally, a {\em uniform matroid} of rank $r$ is a matroid in which any set of size $\leq r$ is independent.

@@

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} Any admissible Wilson loop diagram $W =(\cP, n)$ defines a matroid $M(W)$ with ground set $[n]$. The independent sets are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \label{thm:WLDmatroid}\end{thm}
In other words, the independent sets of $M(W)$ correspond to the sets of vertices in $W$ such that no subset supports fewer propagators than the vertices it contains.

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support different sets of propagators, \bas [n] = \cup_i V(P_i) \quad \textrm{s.t.} \quad \cup P_i = \cP; \quad  P_i \cap P_j = \emptyset \;, \eas the the matroid $M(W)$ is seperable, \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas \note{Closest reference I could find for this was Theorem 3.23 of Agarwala/Amit, which relates to $F(P_i)$ instead of $V(P_i)$; check if there's a better reference?}
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is also in \cite[Lemma FIND]{wilsonloop}.
\end{proof}

@@

\begin{dfn}
  Let $W = (\cP, n)$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

Conversely, suppose we have an exact subdiagram $(P, V(P))$ of $W = (\cP, n)$ supported on $|V(P)| = |P|+3$ vertices, and let $t$ be the subgraph of $\tau(W)$ corresponding to $(P,V(P))$.

@@

% \begin{dfn} \label{restrictiondfn}
% For $W = (\cP, n)$, the restricted diagram, $W|_V$ is the matroid defined by only looking at the vertices $V \subset [n]$.
% \end{dfn}

@@

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram. This gives an algorithm for calculating the set $I_a$, for $a \in [n]$.

@@

If the propagator $p$ is removed at vertex $i$ according to the algorithm for $I_a$, we say that $p$ \emph{contributes} $i$ to $I_a$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram. For each $a \in [n]$, define a function $I_a : \cP \longrightarrow [n]$ by
\[I_a(p) := \text{the vertex label that $p$ contributes to $I_a$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$.
\end{dfn}

@@

%Let $W = (\cP, n)$ be an admissible Wilson loop diagram, and $p = (i,j) \in \cP$, and $a \in [n]$ be a vertex. For $v \in \{i,j\}$, if $P_v(p) \neq \emptyset$, write $P_v(p) = \{s_1 \ldots s_r, t_2, \ldots t_r\}$ as above. Then $p$ does not contribute $v$ to $I_a$ if $i_{s_r} +1 <_{v+1} a <_{v+1} v+1 $. If both $P_v(p)$ and $P_{v+1}(p)$ are non empty, then $p$ also does not contribute $v+1$ to $I_a$.

@@

%Let $W = (\cP, n)$ be an admissible Wilson loop diagram, and $p = (i,j) \in \cP$. Picking a $v \in \{i, j\}$ write $p = (i_p, j_p)$ with $i_p <_{v+1} j_p$. If $a \in [j_p+1, i_p+1]$ then there can be no protecting set $P_v$ or $P_{v+1}$ protecting $p$ at $v$ or $v+1$ from $a$.

@@

%For $W = (\cP, n)$, $p = (i,j)$, and $v \in \{i, j\}$ and $ a \in [n]$ a vertex of $W$, if $v$ (resp. $v+1$) is the first vertex of $V_p$ in the $<_a$ order such that $P_v$ (resp. neither $P_v$ nor $P_{v+1}$) does not exist, then $p$ contributes $v$ (resp. $v+1$) to $I_a$.

@@

%If $W = (\cP, n)$ is an admissible Wilson loop diagram with $|\cP| = k$, then Algorithm~\ref{alg:put GN on WLD} puts exactly $k$ distinct labels on each edge.

@@

%Fix $p = (i, j)$. By symmetry, we need only consider the vertices $j$, $j+1$.
%
%Let $k_1 \ldots k_r$ be the set of propagators of the form $k_x = (i_{k_x}, j)$, with $i <_{j+1} k_1 <_{j+1} \ldots k_r <_{j+1} j$. Write $k_{r+1}$ to be the counterclockwise most propagator with an end on the edge $j-1$, if it exists. If $r \geq 2$ or $r+1 = 2$, then both $P_j(p)$ and $P_{j+1}(p)$ exists for some subset of $a \in [n]$. If $r=1$ or $r+1= 2$ then only $P_j(p)$ exists. Otherwise, $P_j(p) = \emptyset$. 
%
%Consider a protecting set $P_j(p) \subset \cP$ such that $k_1 \in P_j(p)$, if it exists. Write it \bas P_j(p) = \{s_1 \ldots s_r, t_2, \ldots t_r\} \;.\eas Define $b_1 = i_{s_r}+1$. By construction, $P_j(p)$ prevents $p$ from contributing $j$ to $I_a$ for all $a \in [b_1+1, j]$. We see that, if $r >0$, for all $a \in [i+2, b_1]$, $p$ contributes $j$ to $I_a$. (\hlfix{Note that by construction, one cannot have $i_{s_r}+1 <_{j+1} i+2$.}{verify}) This is because in this range, $j$ is the first vertex of $p$ encountered by Algorithm \ref{Alg: put GN on WLD}. Furthermore, by Lemma \ref{protect conv lem}, and definition \ref{protect dfn} if $i_{s_r}+1 \not <_{j+1} i+2$ $P_1$ does not protect $p$ at $j$ for $a$ in this range. If $r = 0$ then $p$ contributes $j$ for all $a \in [i+2, j]$. In this case, we say $b_1 = j$.
%
%Now suppose $r$ is such that both $P_j(p)$ and $P_{j+1}(p)$ exists for some $a$. This time $k_2 \subset P_j(p)$. Define $b_2 = i_{s_r}+1$, where now we use $s_r$ to indicate the last propagator in this new protecting set. By construction, $P_{j+1}(p)$ prevents $p$ from contributing $j+1$ to $I_a$ for all $a \in [b_2+2, j+1]$. Therefore, if both protecting sets exist, for all $a \in [b_1+1, b_2]$ $p$ contributes $j+1$ to $I_a$. This is exactly the range in which there is exactly one valid protecting set. Therefore, in this range, $j+1$ is the first vertex of $p$ without a protective set. If there is no $a$ such that both $P_j(p)$ and $P_{j+1}(p)$ do not hold, then  $p$ contributes $j+1$ for all $a \in [b_1 +1, j+1]$.
%
%Finally, for $r > 1$, and $j_{k_2}=j$,$a \in [b2+1, j+1]$, $p$ contributes $i$ to $I_a$. Again, for all $a$ in this range, both sets $P_j(p)$ and $P_{j+1}(p)$ protect $p$ at $j$ and $j+1$ from $a$. Furthermore, by Corrollary \ref{protect cor}, there can be no protective sets of $i$ or $i+1$ in this range. 
%
%By switching the roles of $i$ and $j$, we see that for every element of $V_p$, there is a non-empty cyclic interval such that $p$ contributes $v$ to all $I_{a}$ for in said interval.

>>>>

%*****************
%Shortcuts for frequently used special characters
\newcommand{\fB}{\mathfrak{B}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\fZ}{\mathfrak{Z}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\Prop}{\textrm{Prop}}
\newcommand{\cW}{\mathcal{W}}
\newcommand{\bM}{\mathbb{M}}
\newcommand{\cZ}{\mathcal{Z}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\Dom}{\textrm{Dom}}
\newcommand{\detzr}[1] {\langle (\cZ_*^\mu|V(p))^{#1} \rangle}
\newcommand{\II}{\mathcal{I}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\BB}{\mathcal{B}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\interval}[2]{[\![#1,#2]\!]}
\newcommand{\gale}[1]{\preccurlyeq_{#1}}
\newcommand{\sgale}[1]{\prec_{#1}}
%*****************

@@

\begin{dfn}\label{WLdfn}
A Wilson loop diagram is given by the following data: a cyclicly ordered set $V$, along with a choice of first vertex (labeled $1$), and $k$ pairs, called propagators, written $\{p_r = (i_r, j_r)\}_{r=1}^k$ with labels ordered such that $i_r +1 < j_r$ relative to the fisrt vertex. \end{dfn}

We depict this data as a circle with marked points, called vertices. The vertices are labeled by $V$ (preserving the cyclic ordering). The arc between consecutive vertices are called edges. There are $k$ wavy lines in the interior of the diagram, depicting the proagators, with endpoints on the edges. A propagator, $p =(i,j)$ has one endpoint on the edge between the vertex labeled $i$ and $i+1$ and another endpoint on the edge defined by $j$ and $j+1$. The condition on $i_r$ and $j_r$ means that the propagator does not go between adjacent edges. Let $\cP = \{p_r\}_{r=1}^k$ be the set of propagators. Then we write \bas W = (\cP, V) \;.\eas

Note that the marked circle gives the vertices of $W$ a cyclic ordering. The choice of first vertex gives it a compatible linear order. Both the cyclic and the linear order become the correct perspective at various points in this paper.

Often we take $V$ to be $[n]$, the cyclically ordered set of integers, $1 \ldots n$. In this case, we write $W = (\cP, [n])$. We introduce some notation to speak of vertices supporting a propagator, and the set of propagators supported on a vertex set.

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, [n])$.
\begin{enumerate}
\item For $p \in \cP$, let $V(p) = \{i_p, i_p+1, j_p, j_p+1\}$ be the set of vertices supporting $p$. Then, for $P \subseteq \cP$, the set $V(P) = \cup_{p \in P} V(p)$ is the vertex support of $P$.
\item For $V \subseteq [n]$, write $\Prop(V) = \{ p \in \cP | V(p) \cap V \neq \emptyset \} $.
\item For $P \subseteq \cP$, define $F(P) = V(P^c)^c$ to be the set of vertices in $[n]$ that only support propagators in the set $P$.
\end{enumerate}
\end{dfn}

Note that $F(\emptyset)$ is the set of vertices that are not in the support of any propagators. \sanote{did we settle on a term here? I'd like to make this a definition}

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram is admissible if \begin{enumerate}
\item $|V| \geq |\cP| + 4$
\item There does not exists a set of propagators, $P \subseteq \cP$ such that $|V(P)| < |P| + 3$.
\item There does not exist a pair of propagators, $p, q \subseteq \cP$ such that $i_p < i_q < j_p <j_q$.
\end{enumerate}
A Wilson loop diagram is weakly admissible if the second and third conditions hold.
 \end{dfn}

The first conditions states that there are at least four more vertices than propagators in an admissible Wilson Loop Diagram. The second imposes an upper bound on how densely the propagators can be fitted in the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram. In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible if and only if $n > \cP +4$, and has neither crossing propagators nor any pairs of propagators that start and end on the same pair of non-adjacent edges.

@@

\begin{dfn} \label{subdiagramdfn}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. The weakly admissible diagram, $W'$ is a subdiagram of $W$, written $W' \subseteq W$, if \bas W' = (P, V); \quad P \subseteq \cP ; \quad V(P) \subseteq V \subseteq [n]\;.\eas
\end{dfn}

@@

\begin{dfn}\label{equivdfn} 
There is an equivalence relationship on the set of admissible Wilson loops diagrams given by the transititve closure the following binary relation: $W = (\cP, [n]) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

Each Wilson loop diagram, $W = (\cP, [n])$ with $|\cP| = k$ is associated to a $k \times n$ matrix with non-zero real variable entries, called $C(W)$:

@@

\begin{dfn}\label{def R(W)}
For $W = (\cP, [n])$, $R(W) = \prod_{e=1}^n R_e$, with $R_e$ defined by the propagators ending on it. For any edge $e$ of $W$, order the propagators incident on $e$ as $\{p_1 \ldots p_r\}$, ordered such that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$. Then \bas R_e =  c_{p_1,e+1} \prod_{j= 1}(r-1) \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_s,e}\;.\eas Note that in this notation, if $r = 1$, $R_e = c_{p,e} c_{p,e+1}$.
\end{dfn}

@@

Finally, we describe several important types of matroids. A matroid of rank $k$ with a ground set of size $n$ is called {\em realizable} if there exists some $A \in Gr(k,n)$ whose non-zero $k\times k$ minors are exactly those with columns indexed by elements of $\cB$. A {\em positroid} is a matroid which can be realized by an element of the totally nonnegative Grassmannian $\Gr(k,n)$. Finally, a {\em uniform matroid} of rank $r$ is a matroid in which any set of size $\leq r$ is independent.

@@

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} Any admissible Wilson loop diagram $W =(\cP, [n])$ defines a matroid $M(W)$ with ground set $[n]$. The independent sets are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \label{thm:WLDmatroid}\end{thm}
In other words, the independent sets of $M(W)$ correspond to the sets of vertices in $W$ such that no subset supports fewer propagators than the vertices it contains.

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support different sets of propagators that form a partition of the propagotor set, \bas [n] = \sqcup_i V(P_i) \quad \textrm{s.t.} \quad \sqcup P_i = \cP; \quad V(P_i) \cap V(P_j) = \emptyset; \quad  P_i \cap P_j = \emptyset \;, \eas then the matroid $M(W)$ is seperable, \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$
\end{proof}

@@

\begin{dfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, [n])$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

Conversely, suppose we have an exact subdiagram $(P, V(P))$ of $W = (\cP, [n])$ supported on $|V(P)| = |P|+3$ vertices, and let $t$ be the subgraph of $\tau(W)$ corresponding to $(P,V(P))$.

@@

% \begin{dfn} \label{restrictiondfn}
% For $W = (\cP, [n])$, the restricted diagram, $W|_V$ is the matroid defined by only looking at the vertices $V \subset [n]$.
% \end{dfn}

@@

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. This gives an algorithm for calculating the set $I_a$, for $a \in [n]$.

@@

If the propagator $p$ is removed at vertex $i$ according to the algorithm for $I_a$, we say that $p$ \emph{contributes} $i$ to $I_a$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. For each $a \in [n]$, define a function $I_a : \cP \longrightarrow [n]$ by
\[I_a(p) := \text{the vertex label that $p$ contributes to $I_a$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$.
\end{dfn}

@@

%Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $p = (i,j) \in \cP$, and $a \in [n]$ be a vertex. For $v \in \{i,j\}$, if $P_v(p) \neq \emptyset$, write $P_v(p) = \{s_1 \ldots s_r, t_2, \ldots t_r\}$ as above. Then $p$ does not contribute $v$ to $I_a$ if $i_{s_r} +1 <_{v+1} a <_{v+1} v+1 $. If both $P_v(p)$ and $P_{v+1}(p)$ are non empty, then $p$ also does not contribute $v+1$ to $I_a$.

@@

%Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $p = (i,j) \in \cP$. Picking a $v \in \{i, j\}$ write $p = (i_p, j_p)$ with $i_p <_{v+1} j_p$. If $a \in [j_p+1, i_p+1]$ then there can be no protecting set $P_v$ or $P_{v+1}$ protecting $p$ at $v$ or $v+1$ from $a$.

@@

%For $W = (\cP, [n])$, $p = (i,j)$, and $v \in \{i, j\}$ and $ a \in [n]$ a vertex of $W$, if $v$ (resp. $v+1$) is the first vertex of $V(p)$ in the $<_a$ order such that $P_v$ (resp. neither $P_v$ nor $P_{v+1}$) does not exist, then $p$ contributes $v$ (resp. $v+1$) to $I_a$.

@@

%If $W = (\cP, [n])$ is an admissible Wilson loop diagram with $|\cP| = k$, then Algorithm~\ref{alg:put GN on WLD} puts exactly $k$ distinct labels on each edge.

@@

%Fix $p = (i, j)$. By symmetry, we need only consider the vertices $j$, $j+1$.
%
%Let $k_1 \ldots k_r$ be the set of propagators of the form $k_x = (i_{k_x}, j)$, with $i <_{j+1} k_1 <_{j+1} \ldots k_r <_{j+1} j$. Write $k_{r+1}$ to be the counterclockwise most propagator with an end on the edge $j-1$, if it exists. If $r \geq 2$ or $r+1 = 2$, then both $P_j(p)$ and $P_{j+1}(p)$ exists for some subset of $a \in [n]$. If $r=1$ or $r+1= 2$ then only $P_j(p)$ exists. Otherwise, $P_j(p) = \emptyset$. 
%
%Consider a protecting set $P_j(p) \subset \cP$ such that $k_1 \in P_j(p)$, if it exists. Write it \bas P_j(p) = \{s_1 \ldots s_r, t_2, \ldots t_r\} \;.\eas Define $b_1 = i_{s_r}+1$. By construction, $P_j(p)$ prevents $p$ from contributing $j$ to $I_a$ for all $a \in [b_1+1, j]$. We see that, if $r >0$, for all $a \in [i+2, b_1]$, $p$ contributes $j$ to $I_a$. (\hlfix{Note that by construction, one cannot have $i_{s_r}+1 <_{j+1} i+2$.}{verify}) This is because in this range, $j$ is the first vertex of $p$ encountered by Algorithm \ref{Alg: put GN on WLD}. Furthermore, by Lemma \ref{protect conv lem}, and definition \ref{protect dfn} if $i_{s_r}+1 \not <_{j+1} i+2$ $P_1$ does not protect $p$ at $j$ for $a$ in this range. If $r = 0$ then $p$ contributes $j$ for all $a \in [i+2, j]$. In this case, we say $b_1 = j$.
%
%Now suppose $r$ is such that both $P_j(p)$ and $P_{j+1}(p)$ exists for some $a$. This time $k_2 \subset P_j(p)$. Define $b_2 = i_{s_r}+1$, where now we use $s_r$ to indicate the last propagator in this new protecting set. By construction, $P_{j+1}(p)$ prevents $p$ from contributing $j+1$ to $I_a$ for all $a \in [b_2+2, j+1]$. Therefore, if both protecting sets exist, for all $a \in [b_1+1, b_2]$ $p$ contributes $j+1$ to $I_a$. This is exactly the range in which there is exactly one valid protecting set. Therefore, in this range, $j+1$ is the first vertex of $p$ without a protective set. If there is no $a$ such that both $P_j(p)$ and $P_{j+1}(p)$ do not hold, then  $p$ contributes $j+1$ for all $a \in [b_1 +1, j+1]$.
%
%Finally, for $r > 1$, and $j_{k_2}=j$,$a \in [b2+1, j+1]$, $p$ contributes $i$ to $I_a$. Again, for all $a$ in this range, both sets $P_j(p)$ and $P_{j+1}(p)$ protect $p$ at $j$ and $j+1$ from $a$. Furthermore, by Corrollary \ref{protect cor}, there can be no protective sets of $i$ or $i+1$ in this range. 
%
%By switching the roles of $i$ and $j$, we see that for every element of $V(p)$, there is a non-empty cyclic interval such that $p$ contributes $v$ to all $I_{a}$ for in said interval.
